EvoSwarm: An Event-based Architecture for Multi-population Optimization Algorithms

### 2020-05-07 draft.tex
REformatting a bit

Next, we show the parameters for each algorithm, with Table \ref{tab:GAparams} for
the GA and Table \ref{tab:PSOparams} for PSO. We obtained these parameters
following the same method as in a previous work \cite{garcia2017benchmarking}.
To obtain the parameters, we tested first on the Rastrigin separable function
with five dimensions. After about fifteen experiments, the most challenging
targets were achieved for this particular function. We tested again with
functions one to three, and after obtaining favorable results, the PSO and GA
algorithm parameters were set. In the GA, we randomly set the mutation and
crossover probabilities to have more heterogeneous workers; % highlight this and add citation - JJ 
% done below - M
in these parameters, the range of values is specified. We did not change these parameters during the 
experiments, and only the population size and number of generations were
provided as parameters. This strategy has been applied successfully for increasing 
the performance of multi-population algorithms without requiring the setting 
of initial parameters \cite{garcia2014randomized}.

\begin{table}
  \small
  \caption{ DEAP GA EvoWorker Parameters }
  \label{tab:GAparams} 
  \centering
  \small
  \begin{tabular}{|l|c|}
    \hline
    Selection & Tournament size=12                            \\ \hline
    Mutation & Gaussian $\mu=0.0$, $\sigma=0.5$, indbp=0.05   \\ \hline
    Mutation Probability & [0.1,0.3]                          \\ \hline
    Crossover & Two Point                                     \\ \hline
    Crossover Probability  & [0.2,0.6]                          \\ \hline
  \end{tabular}
\end{table}

\begin{table}
  \small
  \caption{ EvoloPy PSO Parameters }
  \label{tab:PSOparams} 
  \centering
  \small
  \begin{tabular}{|l|c|}
    \hline
    $V_{max}$ & 6 \\ \hline
    $W_{max}$ & $0.9$ \\ \hline
    $W_{min}$ & $0.2$ \\ \hline
    $C_1$ & 2 \\ \hline
    $C_2$ & 2 \\ \hline
  \end{tabular}
\end{table}

The experiments can be easily reproduced by using the configuration
files the the EvoSwarm application uses to run the noiseless BBOB testbed
experiments as seen in Section~\ref{sec:exeperimentalDefinition}. % Specify this a bit more (with commands such as
             % docker-compose up and so on and use an itemize
             % environment to highlight it and answer to the reviewer
             % - JJ
As a first step, we must deploy the docker application using a
{\tt docker-compose.yml} file, where the number of worker containers is
specified. Then a JSON file containing the configuration parameters of the
experiment has to be provided.  Table \ref{tab:params} shows an example of these
parameters. The {\em GA-PSO Ratio} parameter indicates the proportion of
populations that will use the GA algorithm.  In the example, with a value of
$0.50$ there will be about the same proportion of GA and PSO populations. If we
specify a value of $0$, this will give us an algorithm with only PSO
populations, and finally, a value of  $1$ means that all populations will run
the GA algorithm. % Opening up the door to ask for testing with
                  % different proportion. You have to justify this
                  % value - JJ  
                  % We mention earlier that we replicated the params of another paper - M 
Next, we specify a list indicating which of the 24 benchmark
functions will be tested. In the example, the experiment will use the first five
functions.  The {\em Instances} parameter indicates how many instances of each
function will be tested. {\em Instances}  has a default value of 15. We used this
value because it is the standard for the BBOB benchmark \cite{hansen2016coco}.
In the {\em Dimensions} parameter, we define a list of the dimensions that will
be tested, and we must select additional parameters for each dimension.
According to the maximum number of function evaluations (FEs) we mentioned
earlier, we define for each dimension: the number of populations that will be
generated in the setup, and for each population,  the number of generations, and
population size. Finally, the {\em Iterations} parameter indicates how many
complete loops will be performed. All these parameters give us the maximum
number of FEs that will be performed. For instance, for $D = 2$, the maximum
number of FEs is $200,000$, which is the same as $40*50*10*10$.

\begin{table}
    \small
    \caption{ Experiment configuration example 
    }
    \label{tab:params}
    \centering
    \small
    \begin{tabular}{|l|l|l|l|l|}
      \hline
      \multicolumn{3}{|l|}{Parameter}                    & Type             & Example         \\ \hline
      \multicolumn{3}{|l|}{Worker Containers (Docker Compose)}        & \texttt{int}     & \texttt{8} \\ \hline
      \multicolumn{3}{|l|}{GA-PSO Ratio}                 & \texttt{decimal} & \texttt{0.5}    \\  \hline
      \multicolumn{3}{|l|}{Benchmark Functions}          & \texttt{list}    & \texttt{[1, 2, 3, 4, 5]} \\ \hline
      \multicolumn{3}{|l|}{Instances}                    & \texttt{integer}    & \texttt{15} \\ \hline
      \multicolumn{3}{|l|}{Dimensions}                   & \texttt{list}    & \texttt{[10, 20, 40]}        \\ \hline
      \multicolumn{3}{|l|}{Crossover Probability Range}  & \texttt{list}    & \texttt{[0.2, 0.6]}      \\ \hline
      \multicolumn{3}{|l|}{Mutation  Probability Range}  & \texttt{list}    & \texttt{[0.1, 0.3]}      \\ \hline
      \multicolumn{5}{|c|}{Dimensions}                                                      \\ \hline  
      Dimension               & Generations & Population Size & Populations  &     Iterations    \\ \hline
              10              & 50      & 140                 &      5                 & 30                \\ \hline
              20              & 66      & 200                 &      5                 & 30               \\ \hline
              40              & 80      & 250                 &      5                 & 40                \\ \hline
    \end{tabular}
\end{table}

@@

% What is the main objective of this experiment? Measure scalability?
% - JJ 
% Yes, see below - M
\subsection{Setup}

To evaluate the effect of these parameters on the scalability of the system, we
propose  an experiment in which we selected $f_4$ (Skew Rastrigin-Bueche separable) from the
BBOB testbed. This function has been used because it is computationally demanding,
and in higher dimensions has been difficult for PSO \cite{el2009black} and GAs
\cite{nicolau2009application} to solve with the required FEs. As we
are only comparing in terms of execution time, the results from a single function
can be better understood, since there are fewer factors involved.
Following the procedure described in the previous Section, we executed 
two sets of experiments with five and ten initial populations,
repeating each experiment 30 times using  1, 2, 4, and 8 workers.  We kept the same 
maximum number of FEs, changing the relevant parameters for this. 
See Tables \ref{tab:params:10} and \ref{tab:params:5} 
for the complete list.

\begin{figure}[h!p]
  \centering
  \includegraphics[width=\textwidth]{time_wrkr}
  \caption{time required for the execution of 15 function
  instances, in seconds }

\begin{figure}[h!p]
  \centering
  \includegraphics[width=\textwidth]{evalsworker}
  \caption{Boxplot of the number of evaluations per second, for 30 instances of function $f_4$, with different number of workers and populations sizes }

@@

Another additional issue is related to the algorithm itself. Many more
workers than populations will have a wider range of states of
evolution in the experiment; populations that will be mixed will have
very different fitness states, which means that a part of them will
probably have been not so useful, following the intermediate
disturbance hypothesis. When the number of workers is increased, the
difference between the state of evolution of populations will be
intermediate, leading to better results; eventually, when there are as
many workers as populations, the state will be quite similar, which is
shown with a lower gap in the time won adding more workers.

\begin{table}
    \small
    \caption{Parameters used in the experiments with ten populations
    }
    \label{tab:params:10}
    \vspace{0.25cm}
    \centering
    \small
    \begin{tabular}{|l|c|c|c|c|c|c|}
      \hline
      Dimension        & 2  & 3  & 5  & 10 & 20  & 40  \\ \hline
      Generations      & 40 & 25 & 28 & 50 & 66  & 80  \\ \hline
      Population Size  & 50 & 60 & 60 & 70 & 100 & 125 \\ \hline
      Populations      & 10 & 10 & 10 & 10 & 10  & 10  \\ \hline
      Iterations       & 10 & 20 & 30 & 30 & 30  & 40  \\ \hline  
    \end{tabular}
\end{table}

\begin{table}
    \small
    \caption{Parameters used in the experiments with five populations
    }
    \label{tab:params:5}
    \vspace{0.25cm}
    \centering
    \small
    \begin{tabular}{|l|c|c|c|c|c|c|}
      \hline
      Dimension        & 2  & 3  & 5  & 10 & 20  & 40  \\ \hline
      Generations      & 40 & 25 & 28 & 50 & 66  & 80  \\ \hline
      Population Size  & 100 & 120 & 120 & 140 & 200 & 250 \\ \hline
      Populations      & 5 & 5 & 5 & 5 & 5  & 5  \\ \hline
      Iterations       & 10 & 20 & 30 & 30 & 30  & 40  \\ \hline  
    \end{tabular}
\end{table}

% that's pretty obvious. Say what you intend to prove with the experiments and why - JJ 
% Removed - M
% The effect of a
% different number of populations on execution time is reported in  Section
% \ref{sec:exp1results}.
% The results of the performance comparison between homogeneous and
% heterogeneous search strategies are presented in Section \ref{sec:exp2results}.     
%

\begin{table}[]
    \caption{Speedup by worker and dimension, taking one worker as the baseline.}
    \label{tab:speedup-table}
    \vspace{0.25cm}
    \centering

    \begin{tabular}{c|r|r|r|r|r|r|}
    \cline{2-7}
    \multicolumn{1}{l|}{}           & \multicolumn{3}{c|}{5 Populations}                                          & \multicolumn{3}{c|}{10 Populations}                                         \\ \hline
    \multicolumn{1}{|l|}{Dimension} & \multicolumn{1}{c|}{2W} & \multicolumn{1}{c|}{4W} & \multicolumn{1}{c|}{8W} & \multicolumn{1}{c|}{2W} & \multicolumn{1}{c|}{4W} & \multicolumn{1}{c|}{8W} \\ \hline
    \multicolumn{1}{|c|}{2}         & 1.94                    & 3.12                    & 3.86                    & 1.84                    & 3.38                    & 5.98                    \\ \hline
    \multicolumn{1}{|c|}{3}         & 1.93                    & 2.77                    & 3.15                    & 1.90                    & 3.26                    & 5.96                    \\ \hline
    \multicolumn{1}{|c|}{5}         & 1.91                    & 2.90                    & 3.09                    & 1.91                    & 3.46                    & 6.14                    \\ \hline
    \multicolumn{1}{|c|}{10}        & 1.92                    & 3.03                    & 3.30                    & 1.90                    & 3.61                    & 6.41                    \\ \hline
    \multicolumn{1}{|c|}{20}        & 1.91                    & 3.14                    & 3.27                    & 1.89                    & 3.73                    & 6.71                    \\ \hline
    \multicolumn{1}{|c|}{40}        & 2.00                    & 2.92                    & 3.12                    & 1.95                    & 3.80                    & 7.52                    \\ \hline
    \end{tabular}
\end{table}

\begin{table}[h!tbp]
    \small
    \caption{Boxplot of the total number of evaluations for 30 instances of function $f_4$ , with different number of workers and populations sizes }
      % mark significant differences in some way - JJ

@@

              & 2w      & 20,704.11      & 19,197.58      \\
              & 4w      & 32,563.61      & 35,248.83      \\
              & 8w      & 40,106.09      & 59,744.88      \\
    \hline
    3         & 1w      & 9,599.62       & 9,389.69       \\
              & 2w      & 18,634.43      & 18,064.24      \\
              & 4w      & 27,607.12      & 30,419.51      \\
              & 8w      & 30,189.27      & 56,102.75      \\
    \hline
    5         & 1w      & 9,473.63       & 8,671.92       \\
              & 2w      & 18,181.18      & 16,753.68      \\
              & 4w      & 27,351.11      & 31,145.09      \\
              & 8w      & 29,257.72      & 52,985.83      \\
    \hline
    10        & 1w      & 8,353.71       & 8,145.61       \\
              & 2w      & 15,957.87      & 15,577.79      \\
              & 4w      & 25,696.70      & 29,176.94      \\
              & 8w      & 28,571.52      & 54,348.60      \\
    \hline
    20        & 1w      & 6,920.57       & 6,723.48       \\
              & 2w      & 13,297.58      & 13,136.11      \\
              & 4w      & 22,307.73      & 25,271.93      \\
              & 8w      & 24,179.41      & 46,988.12      \\
    \hline
    40        & 1w      & 5,086.88       & 4,976.05       \\
              & 2w      & 9,896.65       & 9,736.12       \\
              & 4w      & 16,326.54      & 18,695.84      \\
              & 8w      & 19,183.37      & 37,302.60      \\
    \hline  
    \end{tabular}
\end{table}

\begin{table}[]
    \small
    \caption{Speedup achieved with incrementing the number of workers for each dimension showing the p-rate for the Wilcoxon test}
    % mark significant differences in some way - JJ
    \label{tab:speedup:test}
    \vspace{0.25cm}
    \centering
    \begin{tabular}{lllllll}
              &               &               & \multicolumn{2}{l}{5 Populations} & \multicolumn{2}{l}{10 Populations} \\
    Dimension & \multicolumn{2}{l}{Increment} & Speedup         & p-value         & Speedup          & p-value         \\
    2         & 1w            & 2w            & 1.94            & 6E-16           & 1.84             & 2E-16           \\
              & 2w            & 4w            & 1.61            & 2E-15           & 1.83             & 1E-15           \\
              & 4w            & 8w            & 1.24            & 1E-06           & 1.77             & 2E-16           \\
    3         & 1w            & 2w            & 1.93            & 2E-16           & 1.90             & 2E-16           \\
              & 2w            & 4w            & 1.44            & 8E-15           & 1.71             & 2E-16           \\
              & 4w            & 8w            & 1.14            & 3E-05           & 1.83             & 2E-16           \\
    5         & 1w            & 2w            & 1.91            & 9E-16           & 1.91             & 2E-16           \\
              & 2w            & 4w            & 1.52            & 3E-11           & 1.82             & 2E-16           \\
              & 4w            & 8w            & 1.06            & 5E-03           & 1.77             & 2E-16           \\
    10        & 1w            & 2w            & 1.92            & 3E-16           & 1.90             & 2E-16           \\
              & 2w            & 4w            & 1.58            & 3E-16           & 1.89             & 2E-16           \\
              & 4w            & 8w            & 1.09            & 1E-03           & 1.78             & 2E-16           \\
    20        & 1w            & 2w            & 1.91            & 2E-16           & 1.89             & 2E-16           \\
              & 2w            & 4w            & 1.64            & 2E-15           & 1.97             & 2E-16           \\
              & 4w            & 8w            & 1.04            & 1E-02           & 1.80             & 2E-16           \\
    40        & 1w            & 2w            & 2.00            & 2E-16           & 1.95             & 2E-16           \\
              & 2w            & 4w            & 1.46            & 1E-09           & 1.95             & 2E-16           \\
              & 4w            & 8w            & 1.07            & 3E-02           & 1.98             & 2E-16          
    \end{tabular}
    \end{table}

\begin{figure}[h!p]
    \centering
    \includegraphics[width=\textwidth]{evalrate}
    \caption{Boxplot of the number of evaluations per second, for 30 instances of function $f_4$, with different number of workers and populations sizes }
    \label{fig:spworker:ratio}
\end{figure}

>>>>

Next, we show the parameters for each algorithm, with Table \ref{tab:GAparams} for
the GA and Table \ref{tab:PSOparams} for PSO. We obtained these parameters
following the same method as in a previous work \cite{garcia2017benchmarking}.
To obtain the parameters, we tested first on the Rastrigin separable function
with five dimensions. After about fifteen experiments, the most challenging
targets were achieved for this particular function. We tested again with
functions one to three, and after obtaining favorable results, the PSO and GA
algorithm parameters were set. In the GA, we randomly set the mutation and
crossover probabilities to have more heterogeneous workers; % highlight this and add citation - JJ 
% done below - M
in these parameters, the range of values is specified. We did not change these parameters during the 
experiments, and only the population size and number of generations were
provided as parameters. This strategy has been applied successfully for increasing 
the performance of multi-population algorithms without requiring the setting 
of initial parameters \cite{garcia2014randomized}.
%
\begin{table}[h!tbp]
  \small
  \caption{DEAP GA EvoWorker Parameters }
  \label{tab:GAparams} 
  \centering
  \small
  \begin{tabular}{|l|c|}
    \hline
    Selection & Tournament size=12                            \\ \hline
    Mutation & Gaussian $\mu=0.0$, $\sigma=0.5$, indbp=0.05   \\ \hline
    Mutation Probability & [0.1,0.3]                          \\ \hline
    Crossover & Two Point                                     \\ \hline
    Crossover Probability  & [0.2,0.6]                          \\ \hline
  \end{tabular}
\end{table}
%
\begin{table}[h!tbp]
  \small
  \caption{ EvoloPy PSO Parameters }
  \label{tab:PSOparams} 
  \centering
  \small
  \begin{tabular}{|l|c|}
    \hline
    $V_{max}$ & 6 \\ \hline
    $W_{max}$ & $0.9$ \\ \hline
    $W_{min}$ & $0.2$ \\ \hline
    $C_1$ & 2 \\ \hline
    $C_2$ & 2 \\ \hline
  \end{tabular}
\end{table}

The experiments can be easily reproduced by using the configuration
files the the EvoSwarm application uses to run the noiseless BBOB testbed
experiments as seen in Section~\ref{sec:exeperimentalDefinition}. % Specify this a bit more (with commands such as
             % docker-compose up and so on and use an itemize
             % environment to highlight it and answer to the reviewer
             % - JJ
As a first step, we must deploy the docker application using a
{\tt docker-compose.yml} file, where the number of worker containers is
specified. Then a JSON file containing the configuration parameters of the
experiment has to be provided.  Table \ref{tab:params} shows an example of these
parameters. The {\em GA-PSO Ratio} parameter indicates the proportion of
populations that will use the GA algorithm.  In the example, with a value of
$0.50$ there will be about the same proportion of GA and PSO populations. If we
specify a value of $0$, this will give us an algorithm with only PSO
populations, and finally, a value of  $1$ means that all populations will run
the GA algorithm. % Opening up the door to ask for testing with
                  % different proportion. You have to justify this
                  % value - JJ  
                  % We mention earlier that we replicated the params of another paper - M 
Next, we specify a list indicating which of the 24 benchmark
functions will be tested. In the example, the experiment will use the first five
functions.  The {\em Instances} parameter indicates how many instances of each
function will be tested. {\em Instances}  has a default value of 15. We used this
value because it is the standard for the BBOB benchmark \cite{hansen2016coco}.
In the {\em Dimensions} parameter, we define a list of the dimensions that will
be tested, and we must select additional parameters for each dimension.
According to the maximum number of function evaluations (FEs) we mentioned
earlier, we define for each dimension: the number of populations that will be
generated in the setup, and for each population,  the number of generations, and
population size. Finally, the {\em Iterations} parameter indicates how many
complete loops will be performed. All these parameters give us the maximum
number of FEs that will be performed. For instance, for $D = 2$, the maximum
number of FEs is $200,000$, which is the same as $40*50*10*10$.
%
\begin{table}[h!tbp]
    \small
    \caption{ Experiment configuration example 
    }
    \label{tab:params}
    \centering
    \small
    \begin{tabular}{|l|l|l|l|l|}
      \hline
      \multicolumn{3}{|l|}{Parameter}                    & Type             & Example         \\ \hline
      \multicolumn{3}{|l|}{Worker Containers (Docker Compose)}        & \texttt{int}     & \texttt{8} \\ \hline
      \multicolumn{3}{|l|}{GA-PSO Ratio}                 & \texttt{decimal} & \texttt{0.5}    \\  \hline
      \multicolumn{3}{|l|}{Benchmark Functions}          & \texttt{list}    & \texttt{[1, 2, 3, 4, 5]} \\ \hline
      \multicolumn{3}{|l|}{Instances}                    & \texttt{integer}    & \texttt{15} \\ \hline
      \multicolumn{3}{|l|}{Dimensions}                   & \texttt{list}    & \texttt{[10, 20, 40]}        \\ \hline
      \multicolumn{3}{|l|}{Crossover Probability Range}  & \texttt{list}    & \texttt{[0.2, 0.6]}      \\ \hline
      \multicolumn{3}{|l|}{Mutation  Probability Range}  & \texttt{list}    & \texttt{[0.1, 0.3]}      \\ \hline
      \multicolumn{5}{|c|}{Dimensions}                                                      \\ \hline  
      Dimension               & Generations & Population Size & Populations  &     Iterations    \\ \hline
              10              & 50      & 140                 &      5                 & 30                \\ \hline
              20              & 66      & 200                 &      5                 & 30               \\ \hline
              40              & 80      & 250                 &      5                 & 40                \\ \hline
    \end{tabular}
\end{table}

@@

To evaluate the effect of these parameters on the scalability of the system, we
propose  an experiment in which we selected $f_4$ (Skew Rastrigin-Bueche separable) from the
BBOB testbed. This function has been used because it is computationally demanding,
and in higher dimensions has been difficult for PSO \cite{el2009black} and GAs
\cite{nicolau2009application} to solve with the required FEs. As we
are only comparing in terms of execution time, the results from a single function
can be better understood, since there are fewer factors involved.
Following the procedure described in the previous Section, we executed 
two sets of experiments with five and ten initial populations,
repeating each experiment 30 times using  1, 2, 4, and 8 workers.  We kept the same 
maximum number of FEs, changing the relevant parameters for this. 
See Tables \ref{tab:params:10} and \ref{tab:params:5} 
for the complete list. 
%
\begin{figure}[h!tbp]
  \centering
  \includegraphics[width=\textwidth]{time_wrkr}
  \caption{Time required for the execution of 15 function
  instances, in seconds }
  % Which function? - JJ
  \label{fig:spworker}
\end{figure}
%
%
%
\begin{figure}[h!tbp]
  \centering
  \includegraphics[width=\textwidth]{evalsworker}
  \caption{Boxplot of the number of evaluations per second, for 30 instances of function $f_4$, with different number of workers and populations sizes }

@@

Another additional issue is related to the algorithm itself. Many more
workers than populations will have a wider range of states of
evolution in the experiment; populations that will be mixed will have
very different fitness states, which means that a part of them will
probably have been not so useful, following the intermediate
disturbance hypothesis. When the number of workers is increased, the
difference between the state of evolution of populations will be
intermediate, leading to better results; eventually, when there are as
many workers as populations, the state will be quite similar, which is
shown with a smaller gap in the time won adding more workers.
%
%
\begin{table}[h!tbp]
    \small
    \caption{Parameters used in the experiments with ten populations
    }
    \label{tab:params:10}
    \vspace{0.25cm}
    \centering
    \small
    \begin{tabular}{|l|c|c|c|c|c|c|}
      \hline
      Dimension        & 2  & 3  & 5  & 10 & 20  & 40  \\ \hline
      Generations      & 40 & 25 & 28 & 50 & 66  & 80  \\ \hline
      Population Size  & 50 & 60 & 60 & 70 & 100 & 125 \\ \hline
      Populations      & 10 & 10 & 10 & 10 & 10  & 10  \\ \hline
      Iterations       & 10 & 20 & 30 & 30 & 30  & 40  \\ \hline  
    \end{tabular}
\end{table}
%
\begin{table}[h!tbp]
    \small
    \caption{Parameters used in the experiments with five populations
    }
    \label{tab:params:5}
    \vspace{0.25cm}
    \centering
    \small
    \begin{tabular}{|l|c|c|c|c|c|c|}
      \hline
      Dimension        & 2  & 3  & 5  & 10 & 20  & 40  \\ \hline
      Generations      & 40 & 25 & 28 & 50 & 66  & 80  \\ \hline
      Population Size  & 100 & 120 & 120 & 140 & 200 & 250 \\ \hline
      Populations      & 5 & 5 & 5 & 5 & 5  & 5  \\ \hline
      Iterations       & 10 & 20 & 30 & 30 & 30  & 40  \\ \hline  
    \end{tabular}
\end{table}
%
%
% The effect of a
% different number of populations on execution time is reported in  Section
% \ref{sec:exp1results}.
% The results of the performance comparison between homogeneous and
% heterogeneous search strategies are presented in Section \ref{sec:exp2results}.     
%
%
\begin{table}[h!tbp]
    \caption{Speedup by worker and dimension, taking one worker as the baseline.}
    \label{tab:speedup-table}
    \vspace{0.25cm}
    \centering

    \begin{tabular}{c|r|r|r|r|r|r|}
    \cline{2-7}
    \multicolumn{1}{l|}{}           & \multicolumn{3}{c|}{5 Populations}                                          & \multicolumn{3}{c|}{10 Populations}                                         \\ \hline
    \multicolumn{1}{|l|}{Dimension} & \multicolumn{1}{c|}{2W} & \multicolumn{1}{c|}{4W} & \multicolumn{1}{c|}{8W} & \multicolumn{1}{c|}{2W} & \multicolumn{1}{c|}{4W} & \multicolumn{1}{c|}{8W} \\ \hline
    \multicolumn{1}{|c|}{2}         & 1.94                    & 3.12                    & 3.86                    & 1.84                    & 3.38                    & 5.98                    \\ \hline
    \multicolumn{1}{|c|}{3}         & 1.93                    & 2.77                    & 3.15                    & 1.90                    & 3.26                    & 5.96                    \\ \hline
    \multicolumn{1}{|c|}{5}         & 1.91                    & 2.90                    & 3.09                    & 1.91                    & 3.46                    & 6.14                    \\ \hline
    \multicolumn{1}{|c|}{10}        & 1.92                    & 3.03                    & 3.30                    & 1.90                    & 3.61                    & 6.41                    \\ \hline
    \multicolumn{1}{|c|}{20}        & 1.91                    & 3.14                    & 3.27                    & 1.89                    & 3.73                    & 6.71                    \\ \hline
    \multicolumn{1}{|c|}{40}        & 2.00                    & 2.92                    & 3.12                    & 1.95                    & 3.80                    & 7.52                    \\ \hline
    \end{tabular}
\end{table}
%
%
\begin{table}[h!tbp]
    \small
    \caption{Boxplot of the total number of evaluations for 30 instances of function $f_4$ , with different number of workers and populations sizes }
      % mark significant differences in some way - JJ

@@

              & 2w      & 20,704.11      & 19,197.58      \\
              & 4w      & 32,563.61      & 35,248.83      \\
              & 8w      & 40,106.09      & 59,744.88      \\
    \hline
    3         & 1w      & 9,599.62       & 9,389.69       \\
              & 2w      & 18,634.43      & 18,064.24      \\
              & 4w      & 27,607.12      & 30,419.51      \\
              & 8w      & 30,189.27      & 56,102.75      \\
    \hline
    5         & 1w      & 9,473.63       & 8,671.92       \\
              & 2w      & 18,181.18      & 16,753.68      \\
              & 4w      & 27,351.11      & 31,145.09      \\
              & 8w      & 29,257.72      & 52,985.83      \\
    \hline
    10        & 1w      & 8,353.71       & 8,145.61       \\
              & 2w      & 15,957.87      & 15,577.79      \\
              & 4w      & 25,696.70      & 29,176.94      \\
              & 8w      & 28,571.52      & 54,348.60      \\
    \hline
    20        & 1w      & 6,920.57       & 6,723.48       \\
              & 2w      & 13,297.58      & 13,136.11      \\
              & 4w      & 22,307.73      & 25,271.93      \\
              & 8w      & 24,179.41      & 46,988.12      \\
    \hline
    40        & 1w      & 5,086.88       & 4,976.05       \\
              & 2w      & 9,896.65       & 9,736.12       \\
              & 4w      & 16,326.54      & 18,695.84      \\
              & 8w      & 19,183.37      & 37,302.60      \\
    \hline  
    \end{tabular}
\end{table}
%
%
\begin{table}[h!tbp]
    \small
    \caption{Speedup achieved with incrementing the number of workers for each dimension showing the p-rate for the Wilcoxon test}
    % mark significant differences in some way - JJ
    \label{tab:speedup:test}
    \vspace{0.25cm}
    \centering
    \begin{tabular}{lllllll}
              &               &               & \multicolumn{2}{l}{5 Populations} & \multicolumn{2}{l}{10 Populations} \\
    Dimension & \multicolumn{2}{l}{Increment} & Speedup         & p-value         & Speedup          & p-value         \\
    2         & 1w            & 2w            & 1.94            & 6E-16           & 1.84             & 2E-16           \\
              & 2w            & 4w            & 1.61            & 2E-15           & 1.83             & 1E-15           \\
              & 4w            & 8w            & 1.24            & 1E-06           & 1.77             & 2E-16           \\
    3         & 1w            & 2w            & 1.93            & 2E-16           & 1.90             & 2E-16           \\
              & 2w            & 4w            & 1.44            & 8E-15           & 1.71             & 2E-16           \\
              & 4w            & 8w            & 1.14            & 3E-05           & 1.83             & 2E-16           \\
    5         & 1w            & 2w            & 1.91            & 9E-16           & 1.91             & 2E-16           \\
              & 2w            & 4w            & 1.52            & 3E-11           & 1.82             & 2E-16           \\
              & 4w            & 8w            & 1.06            & 5E-03           & 1.77             & 2E-16           \\
    10        & 1w            & 2w            & 1.92            & 3E-16           & 1.90             & 2E-16           \\
              & 2w            & 4w            & 1.58            & 3E-16           & 1.89             & 2E-16           \\
              & 4w            & 8w            & 1.09            & 1E-03           & 1.78             & 2E-16           \\
    20        & 1w            & 2w            & 1.91            & 2E-16           & 1.89             & 2E-16           \\
              & 2w            & 4w            & 1.64            & 2E-15           & 1.97             & 2E-16           \\
              & 4w            & 8w            & 1.04            & 1E-02           & 1.80             & 2E-16           \\
    40        & 1w            & 2w            & 2.00            & 2E-16           & 1.95             & 2E-16           \\
              & 2w            & 4w            & 1.46            & 1E-09           & 1.95             & 2E-16           \\
              & 4w            & 8w            & 1.07            & 3E-02           & 1.98             & 2E-16          
    \end{tabular}
    \end{table}
%
%
\begin{figure}[h!tbp]
    \centering
    \includegraphics[width=\textwidth]{evalrate}
    \caption{Boxplot of the number of evaluations per second, for 30 instances of function $f_4$, with different number of workers and populations sizes }
    \label{fig:spworker:ratio}
\end{figure}
